\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-11-08 metatheory-condensed.tex
sneak peek of linkage transformer use

Inheritance and code reuse can already be expressed through projection
of fields out of linkages and their inclusion into new linkages.
To make common patterns of linkage manipulations more convenient,
\TT provides a ``library'' of \emph{linkage transformers}, whose well-formedness judgments
have form $\goodInh{\Gm}{h}{\lsig_1}{\lsig_2}$. The idea is that
applying~$h$ to a linkage of type~$\TyLkg{\lsig_1}$ yields a linkage of
type~$\TyLkg{\lsig_2}$ (\ruleref{tm/inh}).
%Linkage transformers are second-class; the only way to use them is to
%apply them to linkages to form new linkages.
%
Derived families can be modeled as linkage transformers inductively
constructed from the introduction forms $\InhId$, $\InhExt{h}{t}$, $\InhOv{h}{t}$, etc.
\cref{fig:fmltt-selected} shows the beta rule of an exemplary transformer,
\ruleref{tmeq/ov/beta}.
It can be read as follows: given a linkage of form $\LkgAdd{\lkg}{t_1}$,
applying the transformer $\InhOv{h}{t_2}$ overrides the linkage's last
field~$t_1$ with~$t_2$.
$\InhOv{\InhId}{\wcode{\wsig'_{\texttt{tm}}}}$, for instance, can be used
in \cref{fig:stlc-linkage-typing} to form a derived family that
overrides $\wsig_{\texttt{tm}}$ with an extended
signature~$\wsig'_{\texttt{tm}}$.

@@

The definitive version containing all the rules in \TT is stated in
terms of a meta\-language supporting quotient inductive-inductive types
(QIITs)~\cite{nordvall2010inductive,altkap2016}.
QIITs support equality constructors, which facilitate the expression of
conversion rules (e.g., \ruleref{tmeq/sub/id}).
The formalization is available as a supplemental file in Agda syntax,
though it is not checked by Agda because Agda does not support QIITs
natively.
An effort to mechanize the formalization is ongoing, using Agda's
\lsti{REWRITE} pragma to replace the limited uses of equality reflection.
%a \textbf{pen-and-paper} formalization
%assuming the facility of quotient inductive-inductive
%types~\cite{nordvall2010inductive,altkap2016} and extensional equality
%(equality reflection), available as a supplemental file in a
%\textbf{(pseudo-)Agda syntax}. Our formalization is directly following
%Agda syntax and type-checking (equality-checking) is done manually,
%leaving traces in the comments. We are currently working on mechanizing
%it and relying on Agda's rewriting mechanism to replace our limited
%usage of equality reflection.

We also typeset a more complete set of rules than the space allows and
present it in a supplemental appendix.
%
Another supplemental appendix gives a more complete example than
\cref{fig:stlc-linkage-typing} showing how families are modeled and
type-checked using linkages and linkage transformers.

>>>>

Inheritance and code reuse can already be expressed through projection
of fields out of linkages and their inclusion into new linkages.
To make common patterns of linkage manipulations more convenient,
\TT provides a ``library'' of \emph{linkage transformers}, whose well-formedness judgments
have form $\goodInh{\Gm}{h}{\lsig_1}{\lsig_2}$. The idea is that
applying~$h$ to a linkage of type~$\TyLkg{\lsig_1}$ yields a linkage of
type~$\TyLkg{\lsig_2}$ (\ruleref{tm/inh}).
%Linkage transformers are second-class; the only way to use them is to
%apply them to linkages to form new linkages.

Derived families can be modeled as linkage transformers inductively
constructed from the introduction forms
$\InhId$, $\InhExt{h}{t}$, $\InhOv{h}{t}$, $\InhInh{h}$, etc.
\cref{fig:fmltt-selected} shows the beta rule of an exemplary transformer,
\ruleref{tmeq/ov/beta}.
It can be read as follows: given a linkage of form $\LkgAdd{\lkg}{t_1}$,
applying the transformer $\InhOv{h}{t_2}$ overrides the linkage's last
field~$t_1$ with~$t_2$.
For instance, the construction below shows that 
$\InhOv{\InhId}{\wcode{\wsig'_{\texttt{tm}}}}$ is used as the first step
in creating a derived family that
overrides $\wsig_{\texttt{tm}}$ with an extended
signature~$\wsig'_{\texttt{tm}}$.
\vspace{.9ex}

\noindent
\input{stlcfix-linkage-typing.tex}

\noindent
A supplemental appendix gives a more complete example
showing how families are modeled and
type-checked using linkages and linkage transformers.

@@

The definitive version containing all the rules in \TT is stated in
terms of a meta\-language supporting quotient inductive-inductive types
(QIITs)~\cite{nordvall2010inductive,altkap2016}.
QIITs support equality constructors, which facilitate the expression of
conversion rules (e.g., \ruleref{tmeq/sub/id}).
The formalization is available as a supplemental file in Agda syntax,
though it is not checked by Agda because Agda does not support QIITs
natively.
An effort to mechanize the formalization is ongoing, using Agda's
\lsti{REWRITE} pragma to replace the limited uses of equality reflection.
%a \textbf{pen-and-paper} formalization
%assuming the facility of quotient inductive-inductive
%types~\cite{nordvall2010inductive,altkap2016} and extensional equality
%(equality reflection), available as a supplemental file in a
%\textbf{(pseudo-)Agda syntax}. Our formalization is directly following
%Agda syntax and type-checking (equality-checking) is done manually,
%leaving traces in the comments. We are currently working on mechanizing
%it and relying on Agda's rewriting mechanism to replace our limited
%usage of equality reflection. 
%
We also typeset a more complete set of rules than the space allows and
present it in a supplemental appendix.

### 2022-11-08 stlc-linkage-typing.tex
singleton type
tighten fmltt figures a bit
Venn diagram
stlc case study
sneak peek of linkage transformer use

@@

\definecolor{A-color}{HTML}{E7EFC5}
\definecolor{T-color}{HTML}{FFF1D0}
\definecolor{t-color}{HTML}{FAE0E4}

@@

@@

    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t1-c}(0.01,-0.10)(-0.02,0.28)%

@@

\ \lsti!:=!}

@@

@@

@@

\lsti!FRecursion subst!\ \dadada\lsti!.!

@@

    context of the case handler below:%

@@

    \lsti!Case tm_unit :=!\ %

@@

    \Tkg{\lkg_{20}}

@@

}{\TyTkg{\lsig_{20}}}

>>>>

\setlength\tabcolsep{0pt}

@@

@@

\colorlet{Bkg-color-2}{white}

@@

    \tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t1-c}(0.03,-0.10)(-0.02,0.28)%

@@

\ \;\lsti!:=!}

@@

    \ \ \lsti!: tm!

@@

\cellcolor{Bkg-color-2}

@@

\cellcolor{Bkg-color-2}
\lsti!FRecursion subst!\ \;\dadada\lsti!.!

@@

    context (i.e., $A_6$) of the case handler below:%

@@

    \lsti!Case tm_unit :=!\ \;%

@@

    \Tkg{\lkg_{53}}

@@

}{\TyTkg{\lsig_{53}}}
